\LARGE Untangling the link between traits, size and growth rate in plants

### 2015-02-18 ms.tex
Move from short to medium format paper

% THIS FILE TAKEN FROM PANDOC DEMOS PAGE:http://johnmacfarlane.net/pandoc/demos.html,http://johnmacfarlane.net/pandoc/demo/mytemplate.tex

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript

\usepackage{fontspec}
\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}

@@

% \usepackage{natbib}
% \bibliographystyle{abbrvnat}
% \setcitestyle{authoryear,open={(},close={)}}

\usepackage[sort&compress]{natbib}
\citestyle{nature}
\bibliographystyle{ieeetr}

@@

\usepackage[rgb,dvipsnames]{xcolor}
\definecolor{ashgrey}{rgb}{0.7, 0.75, 0.71}
\definecolor{grey}{rgb}{0.5, 0.5, 0.5}

@@

\usepackage{titling}

\title{Untangling the link between traits, size and growth rate in plants}

\author{
{\bf Daniel Falster} \\ \normalsize\emph{Biological Sciences, Macquarie University} \\ \url{daniel.falster@mq.edu.au}\vspace{0.1in}\\
{\bf Rich FitzJohn} \\ \normalsize\emph{Biological Sciences, Macquarie University} \\ \url{richard.fitzjohn@mq.edu.au}\vspace{0.1in}\\
{\bf Joe Wright} \\ \normalsize\emph{Center for Tropical Forest Science, Smithsonian Tropical Research Institute, Panama, Republic of Panama} \\ \url{wrightj@si.edu}\vspace{0.1in}\\ }

\date{}

\begin{document}
\thispagestyle{empty} % avoid header and footer on first page. Put after \maketitile, if using that
{\LARGE \bf \thetitle \vspace{0.2in}}

\theauthor

\section{Abstract}\label{abstract}

Plant growth is the foundation of the terrestrial biosphere, powering
food-webs, the carbon cycle and much of human
activity \citep{raupach_carbon_2010}. Plant species differ in many
functional traits, including aspects of tissue construction and energy
allocation, that drive differences in mechanical strength, biomass
partitioning, and rates of photosynthesis and tissue
turnover\citep{wright_world-wide_2004,chave_towards_2009, westoby_plant_2002}. Yet, it remains unclear how -- and even if --
such traits influence whole-plant growth, with the simple relationships
predicted by existing theory often lacking empirical
support\citep{wright_functional_2010,poorter_are_2008}. Here we
provide a new set of theoretical predictions, validated using 20 years
of growth data\citep{condit_barro_2012,kitajima_leaf_2013}, showing
that the influence of three prominent traits on growth rate is moderated
by plant size, and that the impact of size on the nature of
relationships varies among traits. We extend a widely-used theoretical
model that links growth rate in seedlings with a single leaf
trait\citep{lambers_inherent_1992,wright_cross-species_2000} to
explicitly include influences of size, light environment, and three
other prominent traits. We find the influence of leaf construction cost
(\(\phi\)) on growth is strong in small plants but weakens with size. In
contrast, the influence of stem construction cost (\(\rho\)) peaks at
intermediate sizes, while the influence of height at maturation
(\(h_m\)) strengthens with size. Our theoretical model suggests the
change in trait-growth relationships with size arises due to changes in
biomass partitioning and reproductive allocation. Moreover, we show how
traits moderate plant responses to light environment and also determine
shade tolerance, supporting diverse empirical
results\citep{ruger_functional_2012,poorter_leaf_2006}. By
disentangling the effects of plant size, light environment and traits on
growth rates, our results provide a solid theoretical foundation for
trait ecology and thus provide a platform for understanding growth
across diverse species around the world.

\section{Main text}\label{main-text}

Functional traits capture core differences in the strategies plants use
to generate and invest surplus energy\citep{wright_world-wide_2004, chave_towards_2009,westoby_plant_2002}. Data for prominent
traits now exists for many of the world's 250000 plant
species\citep{cornwell_functional_2014}; yet the effect of these traits
on growth remains poorly understood. Prominent traits such as \(\phi\)
(also known as leaf mass per area) are widely considered as indicators
of growth rate\citep{reich_world-wide_2014}, because cheaper leaf
construction should enable faster canopy
expansion\citep{blackman_compound_1919,westoby_time_2000}. Indeed,
growth rate in seedlings is strongly and negatively correlated with
\(\phi\)\citep{cornelissen_seedling_1996,wright_cross-species_2000}
(Fig. \ref{f-BCI}). Although, it has been suggested that low \(\phi\)
might also give a growth advantage in larger
plants\citep{enquist_general_2007}, an increasing number of empirical
studies have failed to detect any such
effect\citep{wright_functional_2010, @herault_functional_2011,
@poorter_are_2008}. Meanwhile, expectations on how other prominent
traits might influence growth are lacking. The growing discord between
theoretical expectation and empirical result thus indicates a strong
need for a new theoretical framework through which the influence of
traits and size on plant growth can be understood.

Here we extend a widely-used theoretical model that links growth rate in
seedlings to \(\phi\)\citep{lambers_inherent_1992, wright_cross-species_2000}, to explicitly include influences of
size, light environment, and other prominent traits. This extension
allows us to generate novel predictions on the link between four key
traits and plant demography (Table \ref{tab:predictions}). Predictions
are compared against data from both a long-term
plot\citep{condit_barro_2012} and field
experiment\citep{kitajima_leaf_2013} at Barro Colorado Island, Panama.
We use a new method to estimate species potential growth rate,
accounting for size specific effects (Fig. \ref{S-quantile_examples},
see Methods for details). This new method allowed us to extract a more
definite signal of trait effects than in previous
studies\citep{wright_functional_2010}. The results of this new analysis
(Fig. \ref{f-BCI}) align closely with our theoretical predictions (Table
\ref{tab:predictions}) showing how the effect of traits on growth
changes with plant size, with different traits exhibiting distinct
size-dependent effects.

We begin with a standard model for the amount of biomass available for
growth, \(dP / \textrm{d}t\), given by the difference between income
(total photosynthesis) and losses (respiration and turnover) within the
plant\citep{falster_influence_2011}:

\begin{equation}\label{eq:dPdt}
\underbrace{\strut \frac{dP}{\textrm{d}t}}_\textrm{net biomass production}= \underbrace{\strut Y}_\textrm{yield} \big(\underbrace{\strut \frac{m_\textrm{l}}{\phi} \, A}_\textrm{photosynthesis} - \big(\underbrace{\strut \, \frac{m_\textrm{l}}{\phi} \, r_\textrm{l} + \sum_\textrm{i=b,s,r}{m_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)\big) - \big(\underbrace{\strut \sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}\big).
\end{equation}

Here, \(m,r\), and \(k\) refer to the mass, respiration rate, and
turnover rate of different tissues, denoted by subscripts \(l\)=leaves,
\(b\)=bark, \(s\)=sapwood and \(r\)=roots. \(A\) is the assimilation
rate of CO\(_2\) per leaf area and \(Y\) is yield: the fraction of
assimilated carbon fixed in biomass (the remaining fraction being lost
as growth respiration) (see Table \ref{tab:params} for units and
definitions). Photosynthesis is proportional to leaf area,
\(a_\textrm{l} = m_\textrm{l} / \phi\), while total mass of living
tissues
\(m_\textrm{t}=m_\textrm{l}+m_\textrm{b}+m_\textrm{s}+m_\textrm{r}.\)
Eq. \ref{eq:dPdt} assumes that tissues lost via turnover are replaced
before new growth occurs.

Current theory emphasises the relationship between \(\phi\) and relative
growth in mass\citep{enquist_general_2007},
\(\textrm{d}P / \textrm{d}t / m_\textrm{t}\). For seedlings, which are
young and mostly leaf, it is reasonable to ignore all turnover terms as
well as the respiration terms for non-leaf tissues. Net production then
becomes a linear function of leaf area and net photosynthesis per leaf
area (\(A_\textrm{net} = Y(A - r_\textrm{l})\)), making relative growth
rate a linear function of \(\phi\):

\begin{equation}\label{eq:RGR}
\underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}\frac{1}{m_\textrm{t}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{m_\textrm{l}}{m_\textrm{t}}. \end{equation}

Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
relation to \(\phi\)\citep{wright_cross-species_2000}, this
approximation does not directly link to other traits, or to the
variables that are routinely collected for large trees: namely plant
height (\(h\)) and stem cross-sectional area (\(a_\textrm{st}\)) or
diameter \(D\).

To model growth in either height or basal area requires we to account
not just for mass production, but also for the costs of building new
tissues, allocation to reproduction, and architectural layout. In eq.
\ref{eq:RGR}, these terms are missing, whereas in the model presented
here they are made explicit.

Mathematically, height growth can be decomposed into a product of
physiologically relevant terms\citep{falster_influence_2011}:

\begin{equation} \label{eq:dhdt}
\underbrace{\strut\frac{\textrm{d}h}{\textrm{d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}}_\textrm{architecture}
\times \underbrace{\strut\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}}_\textrm{marginal leaf deployment}
\times \underbrace{\strut\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}}_\textrm{allocation to growth}
\times \underbrace{\strut\frac{\textrm{d}P}{\textrm{d}t}}_\textrm{mass production}.
\end{equation}

The first term on the right of eq \ref{eq:dhdt},
\(\textrm{d}h / \textrm{d}a_\textrm{l}\), is the growth in plant height
per unit growth in total leaf area; accounting for the architectural
strategy of the plant. Some species tend to leaf out more than grow
tall, while other species emphasise vertical
extension\citep{poorter_architecture_2006}.

The second term, \(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\),
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures. As such, \(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\)
can itself be expressed as a sum of construction costs per unit leaf
area:

\begin{equation}\label{eq:daldmt}
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}
= \big(\phi
 + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}\big)^{-1}.
\end{equation}

The third term in eq \ref{eq:dhdt},
\(\textrm{d}m_\textrm{t} / \textrm{d}P\), gives the fraction of net
biomass production (eq. \ref{eq:dPdt}) that is allocated to growth
rather than reproduction or storage.

In a similar way, basal area (\(a_\textrm{st}\)) increment can be
expressed as the sum of increments in sapwood, bark \& heartwood areas
(\(a_\textrm{ss}, a_\textrm{sb}, a_\textrm{sh}\) respectively):
\(\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}= \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}\).
Noting that heartwood grows via turnover of sapwood, such that
\(\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} = k_\textrm{s} a_\textrm{ss}\),
we have an equation for basal area growth that conatins many of the same
elements as eq. \ref{eq:dhdt}:

\begin{equation}\label{eq:dast}
\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}=
\left(\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}a_\textrm{l}}\right) \times
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}} \times \frac{\textrm{d}m_\textrm{t}}{\textrm{d}P} \times \frac{\textrm{d}P}{\textrm{d}t} + k_\textrm{s} a_\textrm{ss}.
\end{equation}

Diameter growth is then given by the geometric relationship between stem
diameter (\(D\)) and \(a_\textrm{st}\):

\begin{equation} \label{eq:dDdt}
\frac{\textrm{d}D}{\textrm{d}t}= \sqrt{\frac{\pi}{a_\textrm{st}}} \frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}.
\end{equation}

Eqs. \ref{eq:dhdt}-\ref{eq:dDdt} are mathematically true, and must
therefore hold for any model of plant growth. To make explicit
predictions, however, requires two additional elements: 1) A specific
model describing how the various mass and area terms
(\(m_\textrm{l}, m_\textrm{s}, m_\textrm{b}, m_\textrm{r}, a_\textrm{l}, a_\textrm{s}, a_\textrm{b}\))
vary relative to one another; and 2) For any trade-offs in plant
function to be specified; this is where traits have an influence. These
two components are as follows.

To address the first of these challenges, we use a functional balance
model\citep{yokozawa_foliage_1995,falster_influence_2011} assuming
1) an allometric relation between total leaf area and plant height, and
2) a constant ratio between total leaf area and each of sapwood
cross-sectional area, bark cross-sectional area, and total root mass.
These assumptions, which are well-supported by data (Fig.
\ref{f-assumptions}), lead immediately to an allometric model for plant
size (Table \ref{tab:allometry}, see Supplementary material for
derivations). Substituting from Table \ref{tab:allometry} into eqs.
\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
for plant growth in relation to size.

This new growth model captures the intrinsically size-dependent nature
of plant growth (Table \ref{tab:predictions}a), recovering well-known
empirical patterns\citep{sillett_increasing_2010, king_size-related_2011}. Height growth shows a hump-shaped pattern
with size, first increasing then decreasing. This pattern results from
systematic changes in the four components of eq. \ref{eq:dhdt} with size
(Fig. \ref{f-hump}), including a strong decline in the fraction of plant
that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
and increasing reproductive allocation. In contrast, basal-area growth
continues to increase with size\citep{sillett_increasing_2010,
@stephenson_rate_2014}, due to an increasing influence of stem
turnover. Diameter growth shows a weakly hump shaped
curve\citep{herault_functional_2011}, tapering off slightly at larger
sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
mainly because of increased reproductive allocation in older trees
(Fig?). All growth measures decrease sharply with size when expressed as
relative growth rates\citep{iida_linking_2014}.

We now incorporate the effects of some prominent traits by embedding
explicit trade-offs within this growth model. Variation in each trait
implies benefits and costs. These affect specific elements of the growth
equations presented above (Table \ref{tab:trade-offs}).

By definition, the main effect of increasing \(\phi\) is to decrease the
marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
growth rate. However, to fully account for the influence of \(\phi\) on
growth we must also account for the decrease in leaf turnover that
results from superior leaf construction\citep{wright_world-wide_2004}.
This is achieved by linking the turnover rate of leaf in eq.
\ref{eq:dPdt} to \(\phi\). We use an allometric scaling relationship of
the form \(k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}\), which has been
shown to describe patterns across diverse
ecosystems\citep{wright_world-wide_2004} (Fig. \ref{fS-leaf}).

Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
\ref{eq:dDdt} predict a relationship between \(\phi\) and growth that
changes with plant size (Figs. \ref{f-BCI}, \ref{f-lma_growth_size}).
Decreasing \(\phi\) has two impacts on growth rate. First, lower
\(\phi\) increases marginal leaf deployment
(\(\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{t}\) by economising on
construction costs. Second, lower \(\phi\) decreases net production
(\(\textrm{d}P / \textrm{d}t\)), due to increased leaf turnover. Whether
lower \(\phi\) increases growth thus depends on the relative magnitude
of these two effects. When plants are small the effect on leaf
deployment rate is bigger and so decreasing \(\phi\) increases growth
rate. When plants are large, the influence of \(\phi\) on leaf
deployment rate is diminished, because the costs of building other
supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
\ref{f-mass_fraction}). The net result is that at larger sizes, low
\(\phi\) is no longer advantageous for growth (Fig.
\ref{f-lma_growth_size}d).

As for leaf, cheaper stem construction (lower \(\rho\)) decreases the
cost of deploying a unit of leaf area, and may thereby increase growth
rates. However, the cost of cheap stem construction must also be
accounted for. However, the direct physiological trade-offs of lower
\(\rho\) are less understood than for \(\phi\). One possibility is that
lower \(\rho\) decreases mechanical strength resulting in higher
mortality\citep{chave_towards_2009,wright_functional_2010}. Under
this scenario, decreasing \(\rho\) will always provide a growth
advantage, because the costs of low \(\rho\) are not realised within the
terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
\ref{tab:trade-offs}). In addition, lower \(\rho\) may increase turnover
of sapwood. Mirroring our assumptions for leaves, we let sapwood
turnover decrease with \(\rho\),
i.e.\(k_\textrm{ss}=\alpha_5 \, \rho^{-\beta_5}\), with \(\beta_5 > 1\).
Under this scenario, decreasing \(\rho\) provides an advantage in height
growth only when the benefits of low \(\rho\) outweigh the costs. This
effect is strongest at intermediate sizes. The decline in height growth
at low \(\rho\) with size is more marked than for either stem basal-area
and diameter growth, because high sapwood turnover also contributes to
this growth measure. Thus diameter growth continues, even as net
production approaches zero.

Unlike leaf and stem traits, height at maturation (\(h_m\)) moderates
growth by adjusting the amount of energy invested in growth, i.e.~the
term \(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}\) in eqs.
\ref{eq:dhdt} and \ref{eq:dast}. Greater height at maturation (\(h_m\))
can thus lead to a growth advantage by increasing
\(\frac{\textrm{d}m_\textrm{t}}{\textrm{d}P}\) (Table
\ref{tab:predictions}). At smaller sizes, there is no differentiation
among species, because most species are focusing on growth (Fig.
\ref{f-BCI}). At larger sizes, individuals of some species are
allocating a majority of their surplus energy to growth, leaving a
strong signal in potential growth rate (Fig. \ref{f-BCI}).

Two other important predictions arise directly from linking traits to
growth rate (Table \ref{tab:predictions}c). First, traits moderate the
responsiveness of growth to changes in light environment. This response
arises simply because species with higher potential growth rate have
greater potential plasticity. Recent analyses have shown that species
with low \(rho\) increase growth more substantially with incraeses in
light\citep{ruger_functional_2012}. This is exactly the behaviour
exhibited in our model (Fig. \ref{f-growth_light}), at least for small
to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
of low \(\phi\) are lost (Fig. \ref{f-growth_light}). This results also
agrees with empirical results showing that growth for species with low
\(rho\) actually decreases as size
increases\citep{ruger_functional_2012}. In contrast, growth rate tends
to increase in size for the species with highest
\(\rho\)\citep{ruger_functional_2012}.

Variation in \(\phi\) also moderates the response of growth to changes
in light, with species having the lowest \(rho\) being most responsive.
However, unlike for \(\rho\), the effect appears only for the smallest
size classes. Accordingly, empirical analyses have detected such a link
in small plants but not for plants \textgreater{} 0.01m
dbh\citep{ruger_functional_2012}.

Second, species with low \(\phi\) and low \(\rho\) are predicted to be
less shade tolerant (Fig. \ref{f-wplcp}). At low \(\phi\) (\(\rho\)),
leaf (sapwood) turnover is higher and thus a greater light income is
needed to offset these costs. As previously suggested
\citep{givnish_adaptation_1988}, shade-tolerance also decreases with
height because as size increase, the total amount of energy needed to
offset respiratory and turnover costs in the stem also increases. These
finding matches well known empirical patterns, where both \(\phi\) and
\(\rho\) have been linked to shade tolerance\citep{poorter_leaf_2006, lusk_ontogenetic_2008,osunkoya_light_1996}.

Conclusion,\ldots{}..

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  traits do effect growth rate, but not in the simple linear fashion
  previously anticipated.
\item
  Relationships change with size due to changes in biomass partitioning
  and reproductive allocation.
\item
  explains why LMA does not show relationship to growth among larger
  plants. Does not mean unimportant for growth, just effects
  concentrated at small sizes.
\item
  in fact, explains plasticity in LMA with size, where LMA tends to
  increase with size \citep{thomas_asymptotic_1999;
  @thomas_photosynthetic_2010}
\item
  Framework can extend to other traits (:

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    In the current model, the only effects of the trait seed size
    (\(s\)) are to influence starting size of seedlings and fecundity of
    the mother. Seed size therefore has no direct influence on absolute
    growth. In may studies seed size has been found to be negatively
    correlated with mass-based relative growth rate of seedlings. This
    effect is recovered here because of the intrinsic decrease in RGR
    with size.
  \item
    leaf N:
  \item
    sapwood area to leaf area
  \end{itemize}
\item
  Take home: Don't look for traits as being fast or slow growth, rather
  consider traits as defining potential trajectory across sizes and
  light environments.
\end{itemize}

\section{Methods}\label{methods}

\subsection{Plant growth model}\label{plant-growth-model}

TREE

\subsection{Data analysis}\label{data-analysis}

Seedling data

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  about data
\item
  method for estimating growth rate
\end{itemize}

Long term plot data

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  about data
\item
  method for estimating growth rate
\end{itemize}

Traits

@@

\section{Figures}\label{figures}

\begin{figure}[htbp]
\centering
\includegraphics{figures/BCI_data.pdf}
\caption{\textbf{The relationship between traits and potential growth
rate varies with plant size.} For XXX species from tropical rain forest
in Panama, we estimated the potential growth rate of individual's in
each species at a series of diameters \(D\), indicated along right hand
side. The size of circles in each panel indicates the number of data
points used to estimate potential growth rate. Traits values were
calculated from representative individuals in each species.
\label{f-BCI}}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figs/allometry.pdf}
\caption{\textbf{Key assumptions of a functional balance allometric
model, evaluated using global dataset.} We used the biomass and
allometry database to evaluate model assumptions about \textbf{a,}
scaling of leaf area with plant height, \textbf{b} Scaling of sapwood
area with leaf area, and \textbf{c} scaling of root mass with leaf area.
Each dot is a single plant. Lines show standardised major axis lines
fitted to data from each site, with intensity of shading adjusted
according to strength of the relationship. Colours indicate vegetation
type. Dashed black lines show values expected under functional-balance
assumption (see Supplementary text for details). \label{f-assumptions}}
\end{figure}

@@

@@

\section{Tables}\label{tables}

\begin{table}[h]
\caption{Theoretical predictions on the relationship between traits and demography}

>>>>

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{authblk} % allows for better author affiliations

@@

\usepackage{natbib}
\bibliographystyle{ecol_let}
\setcitestyle{authoryear,open={(},close={)}}

@@

\usepackage[rgb,dvipsnames]{xcolor}
\definecolor{grey}{rgb}{0.5, 0.5, 0.5}

@@

\title{\LARGE Untangling the link between traits, size and growth rate in plants}

\author[1]{Daniel S. Falster}
\author[1]{Richard G. FitzJohn}
\author[2]{Joe Wright}

\affil[1]{{\footnotesize Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia}}
\affil[2]{{\footnotesize Center for Tropical Forest Science, Smithsonian Tropical Research Institute, Panama, Republic of Panama}}
\renewcommand\Authands{ and }
\date{\vspace{-3em}}

\begin{document}

\maketitle
\thispagestyle{empty} % avoid header and footer on first page. Put after \maketitile, if using that

\section*{Abstract}\label{abstract}

\section*{Introduction}\label{introduction}

Functional traits are thought to capture core differences in the strategies
plants use to generate and invest surplus energy
\citep{wright_world-wide_2004,chave_towards_2009,westoby_plant_2002}. Useful
traits are those that
can be easily measured for diverse species and where differences between
species are large compared to intra-specific variation. Moroever, the uptake
of traits reflects the observation that although most plants have same basic
construction, physiological function and resource requirements, species differ
considerably from one another in rates of carbon, nitrogen and water uptake,
growth, mortality and seed production. Data for prominent traits now exists
for many of the world's 250000 plant species \citep{cornwell_functional_2014}.
The effect of some traits of multiple aspects of plant function has also been
well documented, including mechanical strength, biomass partitioning, rates of
photosynthesis and rates of tissue turnover
\citep {wright_world-wide_2004,chave_towards_2009}. As a result, the idea has
taken hold that the growth startegy and ecology of a plant species might usefully
be characterised simply by knowing its traits.

While the influence of plant traits on elements of plant physiologival
function is increasingly understood, attempts at using traits to predict
demographic rates have only met with mixed success. In seedlings, the trait
leaf mass per area (LMA) -- a central part of the leaf economics spectrum --
is tightly related to plant relative growth rate in mass. LMA and its
correlate leaf lifespan have also been linked to height growth rate for small
seedlings and saplings. These early successes prompted researchers to search
for similar relationships in large plants. However, the results showed that in
large plants, LMA was not correlated with diameter growth rate. Meanwhile,
other traits such as wood density showed strong relationships
to growth in large plants, but only weak relationships in seedlings \citep
{castro-diez_stem_1998}. To date each relationship has been quantified mostly for
one or few traits and at a single life-stage. Yet the picture is emerging is that the
link between traits and growth may be modified by plant size; a finding which would
challenge the asummption applied by many that particular ends of the trait spectrum
translate directly into fast or slower growth.

The challenge of interpreting diverse empirical results linking traits to
growth rate -- or not -- would clearly be easier if we had clearl expectations
on what signal one should expect, given our understanding of how traits
infleunce plant function. This is a primary, yet perhaps underappreciated,
role for theory: to lay expectations based on mechistic principles against
which empircal data can be compared. Current empircial results suggest the
effect of traits on growth changes with size and possibly also light
environment, yet current theory says little about how relationships might come
about. A widely-used model suggests mass-based relative growth rate will be a
linear function of LMA, at least for seeldings.  An extension of model
suggests a similar relationship may hold at larger sizes. But this prediction
doesn't empirical results. Meanwhile, theoretical predictions on how other
traits should influence growth are largely absent.

A further problem with existing models is that the effects of traits are all
realised via mass production, whereas the effect of some prominent traits is
most likely on the quality of tissue construction, or the amoutn of mass
allocated to different tissues. There are two problem here. First, measuring mass
production is only practical for small plants that can be harvested. For larger plants,
growth is more often emasured either as diameter or height growth. For example, the
CTFS monitors diameter growth in 49 large permenenat plots worldwide. Establishing
the link between traits, mass growth and diamater growth is thus essential for predicting
the effect of traits on growth across these diverse datasets.

Here we provide a mechanistic framework describing how traits link to
demography, linking to commonly used metrics such as height and diameter
growth, combine this with fresh empirical insight using largest
collation of growth and trait data available, from long-term plot at
BCI, Panama.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Shows why effect of traits changes with size,size dependent effect
  differs among traits.

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    Contradicts widespread assumption that traits relate to growth in
    simple linear way
  \end{itemize}
\item
  Offers framework for interpreting and understanding empirical results
\end{itemize}

\section*{Fresh empirical motivation}\label{fresh-empirical-motivation}

To help motivate our model, we begin with a fresh analysis of growth dynamics
in the long-term plot at Barro Colorado Island in Panam, where the growth
rates for over 100 species have been recorrded over the last 30 years. A
prior analysis by \citet{wright_functional_2010} found a mdoerate effect of
wood density on growth rate, but almsot no effect of LMA, and only a weak
effect of a thrid trait maximum plant height (HMAX).

Fresh empirical analysis highlights how effect of traits on pot growth
changes with size.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Effect of lma stronger than previously thought,
\item
  Bridges from seedling studies to adults
\end{itemize}

Results stronger than previously reported.

\section*{Materials and Methods}\label{materials-and-methods}

\subsection*{Conceptual framework}\label{conceptual-framework}

Builds on 2011 paper - highlight what's new from there.

Decomposition of diameter and height growth

Functional-balance Model of growth

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  four key assumptions.
\item
  simple linear functions non-critical for results of this paper,
\item
  explore using an example model, but any model with basic size effects
  will display same behaviour

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    details about our model ()
  \item
    show argument hold intuiatively
  \end{itemize}
\end{itemize}

Embedding trait-related trade-offs

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  use table with arrows to
\end{itemize}

Connection to earlier work

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  generalised version
\item
  present old model as dp/dt so consistent with ours
\end{itemize}

\subsection*{Analysis}\label{analysis}

\subsection*{Growth data}\label{growth-data}

\section*{Results}\label{results}

\subsection*{Changes in growth rate with
size}\label{changes-in-growth-rate-with-size}

\subsection*{How traits influence growth
rate}\label{how-traits-influence-growth-rate}

\subsubsection*{LMA}\label{lma}

\subsubsection*{Wood density}\label{wood-density}

\subsubsection*{HMax}\label{hmax}

\section*{Discussion}\label{discussion}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  mechanistic model of height and dbh growth which accounts for effects
  of size, light environment and traits

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    need to account for size long recognized, but not straightforward
  \item
    look at growth in wide range of measures: mass, height, dbh
  \end{itemize}
\item
  Compared to previous work, shows effects of traits more about
  allocation than mass production
\end{itemize}

\subsection*{Tree growth is more than just
photosynthesis}\label{tree-growth-is-more-than-just-photosynthesis}

Recently realisation that allocation and turnover major areas of
uncertainty

Model reconciles idea that traits can have negative impact on carbon
budget, but still have positive influence on growth rates

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Enquist 2007: sla and wd increase mass-production, where as in our
  theory effects on mass production are negative. (also contradicts
  earlier model from Enquist 1999, where argued wood density had no
  effect on mass production)
\end{itemize}

Assumes low lma and wd used to reduce costs of building leaf and stem,
ie. total leaf area and stem corss section remains same. In this way
differs to assumptions of recent work by anten\_role\_2010,
larjavaara\_rethinking\_2010.

Model is realisation of several ideas that have been around for some
time.

\subsection*{Towards a global model of tree
growth}\label{towards-a-global-model-of-tree-growth}

Key patterns explained

Data challenges

Reproductive allocation

\subsection*{Detecting trait signals in
data}\label{detecting-trait-signals-in-data}

Need more than simply look for correlation with mean growth rate

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  understand influences of size on trait of interest
\item
  traits define potential growth rate --\textgreater{} how to extract
  this
\item
  ideally start to use mechanistic models
\end{itemize}

\subsection*{Implications for trait-based
approaches}\label{implications-for-trait-based-approaches}

How do comparative people interpret sla now - beyond fast slow.

Think about traits as defining potential trajectory, rather than making
species fast or slow growing per se.

Selection on traits in different parts of lifecycle

Trait plasticity - can explain, no clear advantage of low lma at large
sizes

Framework for generalising about effect of traits

\section*{Conclusion}\label{conclusion}

@@

\section*{Tables}\label{tables}

\textbf{Table 1 - Key equations of growth model}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  key parameters of model
\end{itemize}

\textbf{Table 2 - effect of traits on demography}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  combines previous tables 1 \& 2
\end{itemize}

@@

\section*{Figures}\label{figures}

\begin{figure}[htbp]
\centering
\includegraphics{figures/BCI_data2.pdf}
\caption{\textbf{The relationship between traits and potential growth
rate varies with plant size.} For XXX species from tropical rain forest
in Panama, we estimated the potential growth rate of individual's in
each species at a series of diameters \(D\), indicated along right hand
side. The size of circles in each panel indicates the number of data
points used to estimate potential growth rate. Traits values were
calculated from representative individuals in each species.
\label{f-BCI}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/SI_size_dhdt.pdf}
\caption{\textbf{Conceptual framework for understanding size-dependent
changes in growth rate.} \label{f-conceptual}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figs/allometry.pdf}
\caption{\textbf{Key assumptions of a functional balance and trait
trade-offs model, evaluated using global dataset.} We used the biomass and
allometry database to evaluate model assumptions about \textbf{a,}
scaling of leaf area with plant height, \textbf{b} Scaling of sapwood
area with leaf area, and \textbf{c} scaling of root mass with leaf area.
Each dot is a single plant. Lines show standardised major axis lines
fitted to data from each site, with intensity of shading adjusted
according to strength of the relationship. Colours indicate vegetation
type. Dashed black lines show values expected under functional-balance
assumption (see Supplementary text for details). \label{f-assumptions}}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figures/max_leaf_above.pdf}
\caption{\textbf{The effect of traits on growth changes with size and
light environment.} Traits moderate the responsiveness of growth to changes in light
environmentThis figure needs to show more of key result with respect to to size.
Possibly separate figure. \label{f-shifts}}
\end{figure}

@@
